---
title: "DDSAnalytics - data analysis for predicting employee turnover"
author: "Katon Pang"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(RCurl)
library(tidyverse)
library(caret)
library(class)
library(e1071)

library(jsonlite)
library(GGally)
library(dplyr)
library(stringr)
library(ggpubr)
library(corrr)
library(knitr)
library(kableExtra)
library(scales)
library(multcomp)
library(onewaytests)
```

```{r}
# Set working directory to project folder
setwd("C:/Users/katon/OneDrive/Documents/DS 6306 Doing Data Science/Week 14 and 15 Case Study 2/")

# Getting Data From S3 URL Using RCurl
case <- read.table(textConnection(getURL(
  "https://s3.us-east-2.amazonaws.com/msdsds6306/CaseStudy2-data.csv"
)), sep=",", header=TRUE)
## There are 870 observation with 36 variables.

# Getting test set for Attrition From S3 URL
test_attr <- read.table(textConnection(getURL(
  "https://s3.us-east-2.amazonaws.com/msdsds6306/CaseStudy2CompSet%20No%20Attrition.csv"
)), sep=",", header=TRUE)

# Getting test set for monthly salary From S3 URL
test_salary <- read.table(textConnection(getURL(
  "https://s3.us-east-2.amazonaws.com/msdsds6306/CaseStudy2CompSet%20No%20Salary.csv"
)), sep=",", header=TRUE)


test <- read.table(textConnection(getURL(
  "https://s3.us-east-2.amazonaws.com/msdsds6306/Case2PredictionsRegressEXAMPLE.csv"
)), sep=",", header=TRUE)
```

```{r}
summary(case)
str(case)

## Within the 36 variables, the data consists of both categorical and quantitative variables. 


# There is no missing data in the dataset.
sum(is.na(case))

```

```{r}
# Employee Attrition pct graph
case %>% group_by(Attrition) %>% summarise(Count=n()) %>% mutate(attr_pct=round(prop.table(Count),2) * 100) %>% 
ggplot(aes(x=Attrition, y=attr_pct)) + geom_bar(stat="identity", fill = "royalblue") + 
geom_text(aes(label= sprintf("%.2f%%", attr_pct)),
            vjust=-0.5, size=3) + labs(x="Employee Attrition", y="Percentage", title="Employee Attrition (%)")

```

# Common top reasons of employees attrition:
## 1. Lack of purpose / bordom: JobInvolvement, JobSatisfaction
## 2. Poor Compensation: Monthly Income
## 3. Overworked: OverTime, WorkLifeBalance
## 4. Poor work/life balance: WorkLifeBalance, DistanceFromHome, BusinessTravel
## 5. Bad Managers: RelationshipSatisfication, YearsWithCurrManager

```{r}
## 1
# Attrition by Job Satisfaction
# Total number of employees in each JobSatisfaction
Attr_JB = case %>% select(JobSatisfaction, Attrition) %>% group_by(JobSatisfaction) %>% summarise(count=n())

# Find out the number of each Attrition status for each JobSatisfaction
# Combine the result with Attr_JB data frame
Attr_JB <- cbind(Attr_JB,as.data.frame.matrix(table(case$JobSatisfaction, case$Attrition)))

# Make Attr_JB into long dataset
Attr_JB_long <- gather(Attr_JB, Attrition, Qty, c(Yes,No))
# Add column and calculate the percentage of attrition
Attr_JB_long$pct = Attr_JB_long$Qty / Attr_JB_long$count *100

# Create the barplot
Attr_JB_long %>% ggplot(aes(x=JobSatisfaction, y=pct, fill=Attrition, label=pct)) + geom_col(width=0.8, position='dodge') + geom_text(aes(label= sprintf("%.1f%%", pct)), position=position_dodge(0.8), vjust=-0.5, size=3) + labs(x="Job Satisfaction", y="Percentage", title="Employee Attrition (%) by Job Satisfaction") + theme_minimal() + scale_fill_manual(values = c("royalblue", "salmon"))

```

```{r}
## 1
# Attrition by Job Involvement
# Total number of employees in each JobInvolvement
Attr_JI = case %>% select(JobInvolvement, Attrition) %>% group_by(JobInvolvement) %>% summarise(count=n())

# Find out the number of each Attrition status for each JobInvolvement
# Combine the result with Attr_JI data frame
Attr_JI <- cbind(Attr_JI,as.data.frame.matrix(table(case$JobInvolvement, case$Attrition)))

# Make Attr_JI into long dataset
Attr_JI_long <- gather(Attr_JI, Attrition, Qty, c(Yes,No))
# Add column and calculate the percentage of attrition
Attr_JI_long$pct = Attr_JI_long$Qty / Attr_JI_long$count *100

# Create the barplot
Attr_JI_long %>% ggplot(aes(x=JobInvolvement, y=pct, fill=Attrition, label=pct)) + geom_col(width=0.8, position='dodge') + geom_text(aes(label= sprintf("%.1f%%", pct)), position=position_dodge(0.8), vjust=-0.5, size=3) + labs(x="Job Involvement", y="Percentage", title="Employee Attrition (%) by Job Involvement") + theme_minimal() + scale_fill_manual(values = c("royalblue", "salmon"))

```

```{r}
## 5
# Attrition by Relationship Satisfaction
# Total number of employees in each RelationshipSatisfaction
Attr_RS = case %>% select(RelationshipSatisfaction, Attrition) %>% group_by(RelationshipSatisfaction) %>% summarise(count=n())

# Find out the number of each Attrition status for each RelationshipSatisfaction
# Combine the result with Attr_RS data frame
Attr_RS <- cbind(Attr_RS,as.data.frame.matrix(table(case$RelationshipSatisfaction, case$Attrition)))

# Make Attr_RS into long dataset
Attr_RS_long <- gather(Attr_RS, Attrition, Qty, c(Yes,No))
# Add column and calculate the percentage of attrition
Attr_RS_long$pct = Attr_RS_long$Qty / Attr_RS_long$count *100

# Create the barplot
Attr_RS_long %>% ggplot(aes(x=RelationshipSatisfaction, y=pct, fill=Attrition, label=pct)) + geom_col(width=0.8, position='dodge') + geom_text(aes(label= sprintf("%.1f%%", pct)), position=position_dodge(0.8), vjust=-0.5, size=3) + labs(x="Relationship Satisfaction", y="Percentage", title="Employee Attrition (%) by Relationship Satisfaction") + theme_minimal() + scale_fill_manual(values = c("royalblue", "salmon"))

```

```{r}
## 5
# Boxplot - YearsWithCurrManager vs RelationshipSatisfaction group by attrition
case$fRelationshipSatisfaction = as.factor(case$RelationshipSatisfaction)

case %>% ggplot(aes(x = fRelationshipSatisfaction, y = YearsWithCurrManager, fill = Attrition)) + geom_boxplot()
```

```{r}
## 5
# Boxplot - YearsWithCurrManager vs RelationshipSatisfaction group by attrition
case$WorkLifeBalance = as.factor(case$WorkLifeBalance)

case %>% ggplot(aes(x = MaritalStatus, y = WorkLifeBalance, fill = Attrition)) + geom_boxplot()

case %>% ggplot(aes(x = MaritalStatus, fill = WorkLifeBalance)) + geom_bar()

```


```{r}
# Run a t-test on attrition status means

case %>% ggplot(aes(x = Attrition, y = YearsWithCurrManager, fill = Attrition)) + geom_boxplot(fill = "royalblue", color = "black", width = 0.3)

A_yes = as.data.frame(case$YearsWithCurrManager[case$Attrition == "Yes"])
A_No = as.data.frame(case$YearsWithCurrManager[case$Attrition == "No"])

t.test(x = A_yes, y = A_No, conf.int = .95, alternative = "two.sided")

# Run a t-test on attrition status means

case %>% ggplot(aes(x = Attrition, y = MonthlyIncome, fill = Attrition)) + geom_boxplot(fill = "royalblue", color = "black", width = 0.3)

Income_yes = as.data.frame(case$MonthlyIncome[case$Attrition == "Yes"])
Income_No = as.data.frame(case$MonthlyIncome[case$Attrition == "No"])

t.test(x = Income_yes, y = Income_No, conf.int = .95, alternative = "two.sided")
```

```{r}
# Attrition by Over Time
# Total number of employees based on OverTime
Attr_OT = case %>% select(OverTime, Attrition) %>% group_by(OverTime) %>% summarise(count=n())

# Find out the number of each Attrition based on OverTime
# Combine the result with Attr_OT data frame
Attr_OT <- cbind(Attr_OT,as.data.frame.matrix(table(case$OverTime, case$Attrition)))

# Make Attr_OT into long dataset
Attr_OT_long <- gather(Attr_OT, Attrition, Qty, c(Yes,No))
# Add column and calculate the percentage of attrition
Attr_OT_long$pct = Attr_OT_long$Qty / Attr_OT_long$count *100

# Create the barplot
Attr_OT_long %>% ggplot(aes(x=OverTime, y=pct, fill=Attrition, label=pct)) + geom_col(width=0.8, position='dodge') + geom_text(aes(label= sprintf("%.1f%%", pct)), position=position_dodge(0.8), vjust=-0.5, size=3) + labs(x="Over Time", y="Percentage", title="Employee Attrition (%) by Over Time") + theme_minimal() + scale_fill_manual(values = c("royalblue", "salmon"))

```

```{r}
# Attrition by Work Life Balance
# Total number of employees based on WorkLifeBalance
Attr_WLB = case %>% select(WorkLifeBalance, Attrition) %>% group_by(WorkLifeBalance) %>% summarise(count=n())

# Find out the number of each Attrition status for each WorkLifeBalance
# Combine the result with Attr_WLB data frame
Attr_WLB <- cbind(Attr_WLB,as.data.frame.matrix(table(case$WorkLifeBalance, case$Attrition)))

# Make Attr_RS into long dataset
Attr_WLB_long <- gather(Attr_WLB, Attrition, Qty, c(Yes,No))
# Add column and calculate the percentage of attrition
Attr_WLB_long$pct = Attr_WLB_long$Qty / Attr_WLB_long$count *100

# Create the barplot
Attr_WLB_long %>% ggplot(aes(x=WorkLifeBalance, y=pct, fill=Attrition, label=pct)) + geom_col(width=0.8, position='dodge') + geom_text(aes(label= sprintf("%.1f%%", pct)), position=position_dodge(0.8), vjust=-0.5, size=3) + labs(x="Work Life Balance", y="Percentage", title="Employee Attrition (%) by Work Life Balance") + theme_minimal() + scale_fill_manual(values = c("royalblue", "salmon"))

```

```{r}
# Attrition by Business Travel
# Total number of employees based on BusinessTravel
Attr_BT = case %>% select(BusinessTravel, Attrition) %>% group_by(BusinessTravel) %>% summarise(count=n())

# Find out the number of each Attrition status for each BusinessTravel
# Combine the result with Attr_WLB data frame
Attr_BT <- cbind(Attr_BT,as.data.frame.matrix(table(case$BusinessTravel, case$Attrition)))

# Make Attr_RS into long dataset
Attr_BT_long <- gather(Attr_BT, Attrition, Qty, c(Yes,No))
# Add column and calculate the percentage of attrition
Attr_BT_long$pct = Attr_BT_long$Qty / Attr_BT_long$count *100

# Create the barplot
Attr_BT_long %>% ggplot(aes(x=BusinessTravel, y=pct, fill=Attrition, label=pct)) + geom_col(width=0.8, position='dodge') + geom_text(aes(label= sprintf("%.1f%%", pct)), position=position_dodge(0.8), vjust=-0.5, size=3) + labs(x="Business Travel", y="Percentage", title="Employee Attrition (%) by Business Travel") + theme_minimal() + scale_fill_manual(values = c("royalblue", "salmon"))

```

```{r}
# Age vs. Attrition -- Generation vs. Attrition
case$Generation <- ifelse(case$Age<37,"Millenials",
ifelse(case$Age>=38 & case$Age<53,"Gen X","Baby Boomers"))
case$Generation = as.factor(case$Generation)

# Total number of employees in each generation
Attr_Gen = case %>% select(Generation, Attrition) %>% group_by(Generation) %>% summarise(count=n())

# Find out percentage of Attrition for each generation
Attr_Gen <- cbind(Attr_Gen,as.data.frame.matrix(table(case$Generation, case$Attrition)))

# Make Attr_JB into long dataset
Attr_Gen_long <- gather(Attr_Gen, Attrition, Qty, c(Yes,No))
# Add column and calculate the percentage of attrition
Attr_Gen_long$pct = Attr_Gen_long$Qty / Attr_Gen_long$count *100

# Create the barplot
Attr_Gen_long %>% ggplot(aes(x=Generation, y=pct, fill=Attrition, label=pct)) + geom_col(width=0.8, position='dodge') + geom_text(aes(label= sprintf("%.1f%%", pct)), position=position_dodge(0.8), vjust=-0.5, size=3) + labs(x="Generation", y="Percentage", title="Employee Attrition (%) by Generation") + theme_minimal() + scale_fill_manual(values = c("royalblue", "salmon"))

```

```{r}
# Attrition by JobRole
# Total number of employees based on JobRole
Attr_JR = case %>% select(JobRole, Attrition) %>% group_by(JobRole) %>% summarise(count=n())

# Find out the number of each Attrition status for each JobRole
# Combine the result with Attr_JR data frame
Attr_JR <- cbind(Attr_JR,as.data.frame.matrix(table(case$JobRole, case$Attrition)))

# Make Attr_RS into long dataset
Attr_JR_long <- gather(Attr_JR, Attrition, Qty, c(Yes,No))
# Add column and calculate the percentage of attrition
Attr_JR_long$pct = Attr_JR_long$Qty / Attr_JR_long$count *100

# Create the barplot
Attr_JR_long %>% ggplot(aes(x=JobRole, y=pct, fill=Attrition, label=pct)) + geom_col(width=1, position='dodge') + geom_text(aes(label= sprintf("%.1f%%", pct)), position=position_dodge(1), vjust=-0.5, size=2.5) + labs(x="Job Role", y="Percentage", title="Employee Attrition (%) by Job Role") + theme_minimal() + theme(axis.text.x = element_text(angle = 90, vjust = 0.3, hjust=1)) +  scale_fill_manual(values = c("royalblue", "salmon"))

```

```{r}
# Attrition by Gender
# Total number of employees based on Gender
Attr_G = case %>% select(Gender, Attrition) %>% group_by(Gender) %>% summarise(count=n())

# Find out the number of each Attrition status for each Gender
# Combine the result with Attr_JR data frame
Attr_G <- cbind(Attr_G,as.data.frame.matrix(table(case$Gender, case$Attrition)))

# Make Attr_RS into long dataset
Attr_G_long <- gather(Attr_G, Attrition, Qty, c(Yes,No))
# Add column and calculate the percentage of attrition
Attr_G_long$pct = Attr_G_long$Qty / Attr_G_long$count *100

# Create the barplot
Attr_G_long %>% ggplot(aes(x=Gender, y=pct, fill=Attrition, label=pct)) + geom_col(width=0.5, position='dodge') + geom_text(aes(label= sprintf("%.1f%%", pct)), position=position_dodge(0.5), vjust=-0.5, size=2.5) + labs(x="Gender", y="Percentage", title="Employee Attrition (%) by Gender") + theme_minimal() +scale_fill_manual(values = c("royalblue", "salmon"))

```

```{r}
# Attrition by Marital Status
# Total number of employees based on Marital Status
Attr_MS = case %>% select(MaritalStatus, Attrition) %>% group_by(MaritalStatus) %>% summarise(count=n())

# Find out the number of each Attrition status for each Marital Status
# Combine the result with Attr_MS data frame
Attr_MS <- cbind(Attr_MS,as.data.frame.matrix(table(case$MaritalStatus, case$Attrition)))

# Make Attr_MS into long dataset
Attr_MS_long <- gather(Attr_MS, Attrition, Qty, c(Yes,No))
# Add column and calculate the percentage of attrition
Attr_MS_long$pct = Attr_MS_long$Qty / Attr_MS_long$count *100

# Create the barplot
Attr_MS_long %>% ggplot(aes(x=MaritalStatus, y=pct, fill=Attrition, label=pct)) + geom_col(width=0.5, position='dodge') + geom_text(aes(label= sprintf("%.1f%%", pct)), position=position_dodge(0.5), vjust=-0.5, size=2.5) + labs(x="Marital Status", y="Percentage", title="Employee Attrition (%) by Marital Status") + theme_minimal() + scale_fill_manual(values = c("royalblue", "salmon"))

```

```{r}
# Attrition by Stock Option Level
# Total number of employees based on Stock Option Level
Attr_SO = case %>% select(StockOptionLevel, Attrition) %>% group_by(StockOptionLevel) %>% summarise(count=n())

# Find out the number of each Attrition status for each Stock Option Level
# Combine the result with Attr_SO data frame
Attr_SO <- cbind(Attr_SO,as.data.frame.matrix(table(case$StockOptionLevel, case$Attrition)))

# Make Attr_SO into long dataset
Attr_SO_long <- gather(Attr_SO, Attrition, Qty, c(Yes,No))
# Add column and calculate the percentage of attrition
Attr_SO_long$pct = Attr_SO_long$Qty / Attr_SO_long$count *100

# Create the barplot
Attr_SO_long %>% ggplot(aes(x=StockOptionLevel, y=pct, fill=Attrition, label=pct)) + geom_col(width=0.5, position='dodge') + geom_text(aes(label= sprintf("%.1f%%", pct)), position=position_dodge(0.5), vjust=-0.5, size=2.5) + labs(x="Stock Option Level", y="Percentage", title="Employee Attrition (%) by Stock Option Level") + theme_minimal() + scale_fill_manual(values = c("royalblue", "salmon"))

```

```{r}
case2 = case

# Change columns into numeric
case2$Attrition[case2$Attrition == "Yes"] = 0
case2$Attrition[case2$Attrition == "No"] = 1
case2$Attrition = as.numeric(case2$Attrition)

case2$nAttrition = case2$Attrition
case2$nAttrition[case2$nAttrition == "Yes"] = 0
case2$nAttrition[case2$nAttrition == "No"] = 1
case2$nAttrition = as.numeric(case2$nAttrition)

case2$BusinessTravel[case2$BusinessTravel == "Non-Travel"] = 0
case2$BusinessTravel[case2$BusinessTravel == "Travel_Rarely"] = 1
case2$BusinessTravel[case2$BusinessTravel == "Travel_Frequently"] = 2
case2$BusinessTravel = as.numeric(case2$BusinessTravel)

case2$Department[case2$Department == "Human Resources"] = 0
case2$Department[case2$Department == "Research & Development"] = 1
case2$Department[case2$Department == "Sales"] = 2
case2$Department = as.numeric(case2$Department)

case2$EducationField[case2$EducationField == "Human Resources"] = 0
case2$EducationField[case2$EducationField == "Life Sciences"] = 1
case2$EducationField[case2$EducationField == "Medical"] = 2
case2$EducationField[case2$EducationField == "Marketing"] = 3
case2$EducationField[case2$EducationField == "Technical Degree"] = 4
case2$EducationField[case2$EducationField == "Other"] = 5
case2$EducationField = as.numeric(case2$EducationField)

case2$Gender[case2$Gender == "Female"] = 0
case2$Gender[case2$Gender == "Male"] = 1
case2$Gender = as.numeric(case2$Gender)

case2$JobRole[case2$JobRole == "Healthcare Representative"] = 0
case2$JobRole[case2$JobRole == "Human Resources"] = 1
case2$JobRole[case2$JobRole == "Laboratory Technician"] = 2
case2$JobRole[case2$JobRole == "Manager"] = 3
case2$JobRole[case2$JobRole == "Manufacturing Director"] = 4
case2$JobRole[case2$JobRole == "Research Director"] = 5
case2$JobRole[case2$JobRole == "Research Scientist"] = 6
case2$JobRole[case2$JobRole == "Sales Executive"] = 7
case2$JobRole[case2$JobRole == "Sales Representative"] = 8
case2$JobRole = as.numeric(case2$JobRole)

case2$MaritalStatus[case2$MaritalStatus == "Divorced"] = 0
case2$MaritalStatus[case2$MaritalStatus == "Single"] = 1
case2$MaritalStatus[case2$MaritalStatus == "Married"] = 2
case2$MaritalStatus = as.numeric(case2$MaritalStatus)

case2$OverTime[case2$OverTime == "Yes"] = 0
case2$OverTime[case2$OverTime == "No"] = 1
case2$OverTime = as.numeric(case2$OverTime)

summary(case2)
str(case2)
```

## ID: Identification variable EmployeeCount: All 1’s EmployeeNumber: Identification Variable StandardHours: All 80’s. All Yes's Over18.  YearsAtCompany, YearsSinceLastPromotion: Have max values that are significantly larger than the 3rd quartile value

```{r}
# 1. Remove ID, EmployeeCount, EmployeeNumber, StandardHours
case2$ID <- c()
case2$EmployeeCount <- c()
case2$EmployeeNumber <- c()
case2$Over18 <- c()
case2$StandardHours <-c()

library(ppcor)
df = case2[c('nAttrition', 'DailyRate', 'DistanceFromHome', 'HourlyRate', 'MonthlyIncome', 'MonthlyRate', 'NumCompaniesWorked', 'PercentSalaryHike', 'TotalWorkingYears', 'TrainingTimesLastYear', 'YearsAtCompany', 'YearsInCurrentRole', 'YearsSinceLastPromotion', 'YearsWithCurrManager', 'JobRole')]

# pairwise partial correlations for each pair of variables
parCor = pcor(df, method = "spearman")
# p value of the test
head(parCor$p.value)
# partial correlation coefficient
head(parCor$estimate)

# See Attrition only
attCor = as.data.frame(parCor$estimate[1,])
attCor$factor = abs(attCor$`parCor$estimate[1, ]`)
arrange(attCor, factor)



fit = lm(Attrition~JobInvolvement + MonthlyIncome + OverTime + YearsWithCurrManager, data=case2)
summary(fit)
```

```{r}
case2$Attrition = as.factor(case2$Attrition)


## Method 1 for imbalanced dataset
# Oversampling 
summary(case2$Attrition)
OnlyYes = case2 %>% filter(Attrition == "Yes") # Give me all Yes
# Oversampling of Yes
OnlyYesOver = rbind(OnlyYes,OnlyYes[sample(seq(1,140,1),(730-140),replace = TRUE),])
dim(OnlyYesOver)

# Oversampling of Yes and the rest of the No
OverSamp = rbind(case2 %>% filter(Attrition == "No"), OnlyYesOver)
dim(OverSamp)

# Make sure they are on the same order. Relevel to Yes as positive class
OverSamp$Attrition = as.factor(OverSamp$Attrition)
OverSamp$Attrition = relevel(OverSamp$Attrition, ref = "Yes")
case2$Attrition = as.factor(case2$Attrition)
case2$Attrition = relevel(case2$Attrition, ref = "Yes")
test$Attrition = relevel(test$Attrition, ref = "Yes")


```

```{r}
# NB

# Model
model = naiveBayes(OverSamp[,c('JobInvolvement', 'MonthlyIncome', 'OverTime', 'YearsWithCurrManager')],OverSamp$Attrition)

model = naiveBayes(OverSamp[,c('JobInvolvement', 'MonthlyIncome', 'OverTime')],OverSamp$Attrition)

model = naiveBayes(OverSamp[,c('JobInvolvement', 'MonthlyIncome', 'OverTime', 'JobRole')],OverSamp$Attrition)

model = naiveBayes(OverSamp[,c('JobInvolvement', 'MonthlyIncome', 'YearsSinceLastPromotion', 'JobRole')],OverSamp$Attrition)

model = naiveBayes(OverSamp[,c('JobInvolvement', 'MonthlyIncome', 'OverTime', 'YearsWithCurrManager', 'JobRole')],OverSamp$Attrition)

model = naiveBayes(OverSamp[,c('JobInvolvement', 'StockOptionLevel', 'OverTime', 'MaritalStatus', 'JobRole')],OverSamp$Attrition)

# Only test set
table(predict(model,test[,c('JobInvolvement', 'MonthlyIncome', 'OverTime', 'YearsWithCurrManager')]),test$Attrition)
confusionMatrix(table(predict(model,test[,c('JobInvolvement', 'MonthlyIncome', 'OverTime', 'YearsWithCurrManager')]),
                      test$Attrition), mode = "everything")

# Entire case2
table(predict(model,case2[,c('JobInvolvement', 'MonthlyIncome', 'OverTime', 'YearsWithCurrManager')]),case2$Attrition)
confusionMatrix(table(predict(model,case2[,c('JobInvolvement', 'MonthlyIncome', 'OverTime', 'YearsWithCurrManager')]),
                      case2$Attrition), mode = "everything")

table(predict(model,OverSamp[,c('JobInvolvement', 'MonthlyIncome', 'OverTime', 'YearsWithCurrManager')]),OverSamp$Attrition)
confusionMatrix(table(predict(model,OverSamp[,c('JobInvolvement', 'MonthlyIncome', 'OverTime', 'YearsWithCurrManager')]),
                      OverSamp$Attrition), mode = "everything")


table(predict(model,case2[,c('JobInvolvement', 'MonthlyIncome', 'OverTime')]),case2$Attrition)
confusionMatrix(table(predict(model,case2[,c('JobInvolvement', 'MonthlyIncome', 'OverTime')]),
                      case2$Attrition), mode = "everything")

table(predict(model,case2[,c('JobInvolvement', 'MonthlyIncome', 'OverTime', 'JobRole')]),case2$Attrition)
confusionMatrix(table(predict(model,case2[,c('JobInvolvement', 'MonthlyIncome', 'OverTime', 'JobRole')]),
                      case2$Attrition), mode = "everything")

table(predict(model,case2[,c('JobInvolvement', 'MonthlyIncome', 'YearsSinceLastPromotion', 'JobRole')]),case2$Attrition)
confusionMatrix(table(predict(model,case2[,c('JobInvolvement', 'MonthlyIncome', 'YearsSinceLastPromotion', 'JobRole')]),
                      case2$Attrition), mode = "everything")

table(predict(model,OverSamp[,c('JobInvolvement', 'MonthlyIncome', 'YearsSinceLastPromotion', 'JobRole')]),OverSamp$Attrition)
confusionMatrix(table(predict(model,OverSamp[,c('JobInvolvement', 'MonthlyIncome', 'YearsSinceLastPromotion', 'JobRole')]),
                      OverSamp$Attrition), mode = "everything")

table(predict(model,case2[,c('JobInvolvement', 'MonthlyIncome', 'OverTime', 'YearsWithCurrManager', 'JobRole')]),case2$Attrition)
confusionMatrix(table(predict(model,case2[,c('JobInvolvement', 'MonthlyIncome', 'OverTime', 'YearsWithCurrManager', 'JobRole')]),
                      case2$Attrition), mode = "everything")

table(predict(model,OverSamp[,c('JobInvolvement', 'MonthlyIncome', 'OverTime', 'YearsWithCurrManager', 'JobRole')]),OverSamp$Attrition)
confusionMatrix(table(predict(model,OverSamp[,c('JobInvolvement', 'MonthlyIncome', 'OverTime', 'YearsWithCurrManager', 'JobRole')]),
                      OverSamp$Attrition), mode = "everything")

confusionMatrix(table(predict(model,case2[,c('JobInvolvement', 'StockOptionLevel', 'OverTime', 'MaritalStatus', 'JobRole')]),
                      case2$Attrition), mode = "everything")

confusionMatrix(table(predict(model,OverSamp[,c('JobInvolvement', 'StockOptionLevel', 'OverTime', 'MaritalStatus', 'JobRole')]),
                      OverSamp$Attrition), mode = "everything")

```

```{r}
# kNN
set.seed(123)
trainIndices = sample(seq(1:dim(OverSamp)[1]),
                      round(.7*dim(OverSamp)[1]))
train = OverSamp[trainIndices,]
test = OverSamp[-trainIndices,]
head(train)
head(test)
classifications = knn(train[,c('MonthlyIncome', 'YearsSinceLastPromotion', 'JobRole')],test[,c('MonthlyIncome', 'YearsSinceLastPromotion', 'JobRole')], train$Attrition, prob = TRUE, k = 5)

confusionMatrix(table(classifications,test$Attrition), mode = "everything")

classifications = knn(train[,c('JobInvolvement', 'MonthlyIncome', 'OverTime', 'YearsWithCurrManager', 'JobRole')],test[,c('JobInvolvement', 'MonthlyIncome', 'OverTime', 'YearsWithCurrManager', 'JobRole')], train$Attrition, prob = TRUE, k = 5)

confusionMatrix(table(classifications,test$Attrition), mode = "everything")


# LOOCV

OverSamp$JobRole[OverSamp$JobRole == "Healthcare Representative"] = 0
OverSamp$JobRole[OverSamp$JobRole == "Human Resources"] = 1
OverSamp$JobRole[OverSamp$JobRole == "Laboratory Technician"] = 2
OverSamp$JobRole[OverSamp$JobRole == "Manager"] = 3
OverSamp$JobRole[OverSamp$JobRole == "Manufacturing Director"] = 4
OverSamp$JobRole[OverSamp$JobRole == "Research Director"] = 5
OverSamp$JobRole[OverSamp$JobRole == "Research Scientist"] = 6
OverSamp$JobRole[OverSamp$JobRole == "Sales Executive"] = 7
OverSamp$JobRole[OverSamp$JobRole == "Sales Representative"] = 8
OverSamp$JobRole = as.numeric(OverSamp$JobRole)

OverSamp$MaritalStatus[OverSamp$MaritalStatus == "Divorced"] = 0
OverSamp$MaritalStatus[OverSamp$MaritalStatus == "Single"] = 1
OverSamp$MaritalStatus[OverSamp$MaritalStatus == "Married"] = 2
OverSamp$MaritalStatus = as.numeric(OverSamp$MaritalStatus)

OverSamp$OverTime[OverSamp$OverTime == "Yes"] = 0
OverSamp$OverTime[OverSamp$OverTime == "No"] = 1
OverSamp$OverTime = as.numeric(OverSamp$OverTime)

classifications = knn.cv(OverSamp[,c('MonthlyIncome', 'YearsSinceLastPromotion', 'JobRole')],OverSamp$Attrition, k=5)

confusionMatrix(table(classifications, OverSamp$Attrition), mode = "everything")

classifications = knn.cv(OverSamp[,c('JobInvolvement', 'MonthlyIncome', 'OverTime', 'YearsWithCurrManager', 'JobRole')],OverSamp$Attrition, k=5)

confusionMatrix(table(classifications, OverSamp$Attrition), mode = "everything")

### Chosen Attrition model
classifications = knn.cv(OverSamp[,c('JobInvolvement', 'StockOptionLevel', 'OverTime', 'MaritalStatus', 'JobRole')],OverSamp$Attrition, k=5)

confusionMatrix(table(classifications, OverSamp$Attrition), mode = "everything")

## ?? close to last, more everything except specificity
classifications = knn.cv(OverSamp[,c('JobInvolvement', 'StockOptionLevel', 'OverTime', 'MaritalStatus', 'JobRole', 'YearsWithCurrManager')],OverSamp$Attrition, k=5)

confusionMatrix(table(classifications, OverSamp$Attrition), mode = "everything")

## From cor test
classifications = knn.cv(OverSamp[,c('JobInvolvement', 'TotalWorkingYears', 'OverTime', 'JobLevel', 'YearsInCurrentRole')],OverSamp$Attrition, k=5)

confusionMatrix(table(classifications, OverSamp$Attrition), mode = "everything")
```

```{r}

## Method 2 for imbalanced dataset
# New Threshold
summary(case2$Attrition)
# 16% Yes

# knn
set.seed(123)
trainIndices = sample(seq(1:dim(case2)[1]),
                      round(.7*dim(case2)[1]))
train = case2[trainIndices,]
test = case2[-trainIndices,]
head(train)
head(test)
classifications = knn(train[,c('JobInvolvement', 'MonthlyIncome', 'OverTime', 'YearsWithCurrManager', 'JobRole')],test[,c('JobInvolvement', 'MonthlyIncome', 'OverTime', 'YearsWithCurrManager', 'JobRole')], train$Attrition, prob = TRUE, k = 5)

confusionMatrix(table(classifications, test$Attrition), mode = "everything")

# Look at the attributes (Note that probabiliities are all greater than .5)
classifications # classifcaiton with probabilities
attributes(classifications) # Look at possible attributes
attributes(classifications)$prob # Probability of what was classified for that observation

#Get probs of a FRAUD specifically
probs = ifelse(classifications == "Yes",attributes(classifications)$prob, 1- attributes(classifications)$prob)

NewClass = ifelse(probs > .15, "Yes", "No")
table(NewClass,test[,2])

NewClass$Attrition = relevel(case2$Attrition, ref = "Yes")
test$Attrition = relevel(test$Attrition, ref = "Yes")

CM_F = confusionMatrix(table(relevel(as.factor(NewClass), ref = "Yes"),relevel(test[,2],ref = "Yes")), mode = "everything")
CM_F

Macro_F1_New_Thresh = mean(c(CM_F[4]$byClass["F1"],CM_NF[4]$byClass["F1"]))
Macro_F1_New_Thresh

```


```{r}
# Run as factor
case$fAttrition = as.factor(case$Attrition)
case$BusinessTravel = as.factor(case$BusinessTravel)
case$Department = as.factor(case$Department)
case$EducationField = as.factor(case$EducationField)
case$Gender = as.factor(case$Gender)
case$JobRole = as.factor(case$JobRole)
case$MaritalStatus = as.factor(case$MaritalStatus)
case$OverTime = as.factor(case$OverTime)
```



```{r}
case %>% select(Attrition, JobSatisfaction, JobInvolvement) %>% ggpairs(aes(color = Attrition))

case %>% select(Attrition, MonthlyIncome, JobInvolvement) %>% ggpairs(aes(color = Attrition))

case %>% select(Attrition, RelationshipSatisfaction, JobInvolvement) %>% ggpairs(aes(color = Attrition))

case %>% select(Attrition, RelationshipSatisfaction, YearsWithCurrManager) %>% ggpairs(aes(color = Attrition))

case %>% filter(JobInvolvement == 1) %>% group_by(Attrition) %>% summarise(count=n()) 


case %>% select(fAttrition, Age, NumCompaniesWorked) %>% ggpairs(aes(color = fAttrition))
case %>% select(fAttrition, JobInvolvement, JobSatisfaction) %>% ggpairs(aes(color = fAttrition))
case %>% select(fAttrition, Age) %>% ggpairs(aes(color = fAttrition))
cor.test(case$JobInvolvement, case$JobSatisfaction)
```
